---
title: "narwhal_habitat_model"
author: "Marie Zahn"
date: '2022-12-12'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/marie/Documents/OMG_Narwhals/OMG_narwhals_habitat-model')
```

```{r load packages}
library(tidyverse)
library(corrplot)
library(MuMIn)
library(arm)
library(faraway)
library(ggplot2)
library(sjPlot)
library(here)
```

Open Data
```{r open rdata files}
# reload data from anywhere
load(here('data-cleaning_R/narwhal_model_data.rdata')) # only observations for open-water period (SI < 75%)
load(here('data-cleaning_R/narwhal_model_data_whales.rdata')) # only observations for when narwhals were in the area
```

## Compare detections in 2019

```{r narwhal detection totals}
## pool soundtrap and aural detections for 2019 data (i.e., take only one obs/day for each site)
mdl_data_open_water_pooled_2019 <- mdl_data_open_water %>% group_by(site,time) %>% 
  filter(year==2019, narwhal==max(narwhal),(row_number()==n()))
## pool detections for entire dataset (i.e., pool 2019 but retain 2018 detections)
mdl_data_open_water_pooled <- mdl_data_open_water %>% group_by(site,time) %>% 
  filter(narwhal==max(narwhal),(row_number()==n())) %>% dplyr::select(-device)
## pool detections for 'whale' data (=only when whales were in the region)
mdl_data_whales_pooled <- mdl_data_whales %>% group_by(site,time) %>% 
  filter(narwhal==max(narwhal),(row_number()==n())) %>% dplyr::select(-device)

## identify how many detections from the same day occurred at more than one site
length(unique(mdl_data_open_water_pooled$time)) # number of days (obs) for open water period
sum(mdl_data_open_water_pooled$narwhal) # total of 93 detections over two years at 3 sites
narwhal_dets <- mdl_data_open_water_pooled %>% filter(narwhal==1) # number of days that narwhals were detected
length(unique(narwhal_dets$time)) # number of unique days when narwhals were identified

## determine how many days whales were seen at more than one glacier front
mdl_data_open_water_pooled %>% dplyr::select(time, site, narwhal) %>% pivot_wider(names_from = site, values_from = narwhal) %>% mutate(rowsums = sum(kong,rink,sver,na.rm=TRUE)) %>% filter(rowsums>1)

## tally up total detections for each site for a given year
rink_2018 <- mdl_data_open_water_pooled %>% filter(year==2018, site=='rink')
rink_2019 <- mdl_data_open_water_pooled %>% filter(year==2019, site=='rink')
sum(rink_2018$narwhal)
sum(rink_2019$narwhal)

kong_2018 <- mdl_data_open_water_pooled %>% filter(year==2018, site=='kong')
kong_2019 <- mdl_data_open_water_pooled %>% filter(year==2019, site=='kong')
sum(kong_2018$narwhal)
sum(kong_2019$narwhal)

sver_2018 <- mdl_data_open_water_pooled %>% filter(year==2018, site=='sver')
sum(sver_2018$narwhal)

```

Out of the 77 days that narwhals were detected, 16 of those days were occasions when whales were detected at two locations on the same day. Whales were never detected at three locations on the same day in 2018.

Totals between each year/site

2018
* Rink: 13
* Kong Oscar: 8
* Sverdrup: 7

2019
* Rink: 32
* Kong Oscar: 33


```{r aural vs soundtrap 2019}
## see whether there are any aural detections that were NOT detected by the soundtrap in 2019
aural_2019 <- mdl_data_open_water %>% group_by(site) %>% filter(year==2019,device=='aural') %>% dplyr::select(time,narwhal) %>% rename(narwhal_aural=narwhal)
sound_2019 <- mdl_data_open_water %>% group_by(site) %>%  filter(year==2019,device=='soundtrap') %>% dplyr::select(time,narwhal)%>% rename(narwhal_sound=narwhal)

compare <- aural_2019 %>% group_by(site) %>%  left_join(sound_2019,by="time") %>% filter(site.x==site.y, narwhal_aural==1 & narwhal_sound==0)
compare # zero rows = no aural detections that were not picked up by soundtrap

## check sums across years
sum(mdl_data_open_water_pooled$narwhal)
mdl_data_open_water %>% filter(year==2018) %>% dplyr::select(narwhal) %>% sum()
mdl_data_open_water %>% filter(year==2019,device=='soundtrap') %>% dplyr::select(narwhal) %>% sum()

## double check comparison using different method
aural_2019_kong <- aural_2019 %>% filter(site=='kong') 
sound_2019_kong <- sound_2019 %>% filter(site=='kong') 
aural_2019_rink <- aural_2019 %>% filter(site=='rink') 
sound_2019_rink <- sound_2019 %>% filter(site=='rink') 

aural_2019_kong %>% left_join(sound_2019_kong,by="time") %>% filter(narwhal_aural==1&narwhal_sound==0)
aural_2019_rink %>% left_join(sound_2019_rink,by="time") %>% filter(narwhal_aural==1&narwhal_sound==0)

```


```{r normalize glacier length data}
## normalize at the start of each year
ref <- mdl_data_open_water_pooled %>% group_by(year, site) %>% filter(time==min(as.Date(time))) %>% dplyr::select(site,year,glacier_length) %>% dplyr::rename(length_ref=glacier_length)

## add normalized glacier length variable and drop unneeded variables
mdl_data_pooled <- mdl_data_open_water_pooled %>% left_join(ref,by=c("site","year")) %>% mutate(glacier_len_norm = glacier_length - length_ref) %>% dplyr::select(!c(length_ref,noise_100Hz,noise_2kHz,runoff_mar,glacier_length))

## sanity check plot
mdl_data_open_water_pooled %>% ggplot(aes(x=as.Date(time),y=glacier_length,color=site))+
  geom_point()
mdl_data_pooled %>% ggplot(aes(x=as.Date(time),y=glacier_len_norm,color=site))+
  geom_point()

## do the same for the other dataset: when only whales were present
ref <- mdl_data_whales_pooled %>% group_by(year, site) %>% filter(time==min(as.Date(time))) %>% dplyr::select(site,year,glacier_length) %>% dplyr::rename(length_ref=glacier_length)

mdl_data_pooled_whales <- mdl_data_whales_pooled %>% left_join(ref,by=c("site","year")) %>% mutate(glacier_len_norm = glacier_length - length_ref) %>% dplyr::select(!c(length_ref,noise_100Hz,noise_2kHz,runoff_mar,glacier_length))

```

```{r check correlations}
## check correlations (again) now that glacier length has been normalized

## open water period
mdl_data_corr <- mdl_data_pooled %>% ungroup() %>% dplyr::select(!c(narwhal, time, site, year)) %>% cor(use = "complete.obs")
## plot correlation test
corrplot::corrplot(mdl_data_corr, method = 'number')

## Hide upper triangle and export csv of correlations
mdl_data_pooled_tbl <- mdl_data_pooled %>% ungroup() %>% dplyr::select(!c(narwhal, time, site, year)) %>% cor(use = "complete.obs") %>% round(2) %>% as.data.frame()

mdl_data_pooled_tbl[upper.tri(mdl_data_pooled_tbl)]<-""
mdl_data_pooled_tbl<-as.data.frame(mdl_data_pooled_tbl)
write.csv(mdl_data_pooled_tbl, here("analysis/mdl_data_pooled_correlations.csv"),row.names = TRUE)

## salt_surf, salt_deep, and temp_deep all correlated with one another and with DOY
## keep DOY and temp_surf
mdl_data_corr <- mdl_data_pooled %>% ungroup() %>% dplyr::select(!c(narwhal, time, site, year, velocity, salt_surf, salt_deep, temp_deep)) %>% cor(use = "complete.obs")

## plot correlation test
corrplot::corrplot(mdl_data_corr, method = 'number')

## -----
## do the same for the subset of data for when whales were occupying Melville Bay
mdl_data_corr <- mdl_data_pooled_whales %>% ungroup() %>% dplyr::select(!c(narwhal, time, site, year)) %>% cor(use = "complete.obs")

# Hide upper triangle and export csv of correlations
mdl_data_pooled_whales_tbl <- mdl_data_pooled_whales %>% ungroup() %>% dplyr::select(!c(narwhal, time, site, year)) %>% cor(use = "complete.obs") %>% round(2) %>% as.data.frame()

mdl_data_pooled_whales_tbl[upper.tri(mdl_data_pooled_whales_tbl)]<-""
mdl_data_pooled_whales_tbl<-as.data.frame(mdl_data_pooled_whales_tbl)
write.csv(mdl_data_pooled_whales_tbl, here("analysis/mdl_data_pooled_whales_correlations.csv"),row.names = TRUE)

```


## Run models

```{r global glm | open water}
## logistic regression 
## glm with binomial response and logit link 

## make sure year is coded as a factor
mdl_data_pooled <- mdl_data_pooled %>% mutate_at(vars(year), list(factor))

## run model
mdl_global_open_water <- glm(narwhal ~ 
                               temp_surf+I(temp_surf^2)+
                               ice_cover_percent+
                               runoff_racmo+
                               velocity+
                               glacier_len_norm+
                               noise_4kHz+
                               year+
                               DOY+I(DOY^2)+
                               site,
                             data = mdl_data_pooled,
                             family=binomial(link="logit"),
                             na.action="na.fail")

# print summary output
summary(mdl_global_open_water)
faraway::sumary(mdl_global_open_water)
# run all possible model combinations
dredge_mdl_open_water <- dredge(mdl_global_open_water)
head(dredge_mdl_open_water)
# export table to csv
write.csv(as.data.frame(dredge_mdl_open_water), row.names = FALSE,
          here("analysis/csv_outputs/mdl_open-water_results.csv"))

```

```{r best model - open water}
mdl_best_open_water <- glm(narwhal ~
                             ice_cover_percent+
                             DOY+I(DOY^2)+
                             year,
                           data = mdl_data_pooled,
                           family=binomial(link="logit"),
                           na.action="na.fail")

# print summary output
summary(mdl_best_open_water)
faraway::sumary(mdl_best_open_water)

# plot predictions
plot_model(mdl_best_open_water)
plot_model(mdl_best_open_water, type = 'pred',
           transform = "exp", show.data=TRUE)

# reference this website: http://www.amelia.mn/sds291/labs/lab_logistic_regression.html
```

```{r}
## compute R^2
# deviances
nn <- 467
DM_mod <-mdl_best_open_water$deviance
D0_mod <- mdl_best_open_water$null.deviance
# R^2
(R2 <- (1 - exp((DM_mod - D0_mod) / nn)) / 
    (1 - exp(-D0_mod / nn)))
```


```{r plot logistic regression result with observations}
## sea ice cover ---------------------------------------
## plot narwhal prob vs sea ice with confidence intervals
clr <- viridis::plasma(1, 0.8, 0.5, 0.5)
## get fitted values
newdata18 <- data.frame(ice_cover_percent = seq(0, 75),
                      DOY = rep(mean(mdl_data_pooled$DOY),length(seq(0, 75))),
                      year = as.factor(rep(2018,length(seq(0, 75)))))
newdata19 <- data.frame(ice_cover_percent = seq(0, 75),
                      DOY = rep(mean(mdl_data_pooled$DOY),length(seq(0, 75))),
                      year = as.factor(rep(2019,length(seq(0, 75)))))

eta18 <- predict(mdl_best_open_water, newdata18)
eta19 <- predict(mdl_best_open_water, newdata19)
# transform out of logit space
p_hat18 <- 1 / (1 + exp(-eta18))
p_hat19 <- 1 / (1 + exp(-eta19))

## get the SE of the response
se18 <- predict(mdl_best_open_water, newdata18, type = "link", se.fit = TRUE)$se.fit
CI_upper18 <- 1 / (1 + exp(-(eta18 + 1.96*se18))) ## upper 95% CI
CI_lower18 <- 1 / (1 + exp(-(eta18 - 1.96*se18))) ## lower 95% CI

se19 <- predict(mdl_best_open_water, newdata19, type = "link", se.fit = TRUE)$se.fit
CI_upper19 <- 1 / (1 + exp(-(eta19 + 1.96*se19))) ## upper 95% CI
CI_lower19 <- 1 / (1 + exp(-(eta19 - 1.96*se19))) ## lower 95% CI

## 95% CI for beta_i based on profile likelihood
round(confint(mdl_best_open_water), 2)

## set plot area
par(mai = c(0.9, 0.9, 0.1, 0.1),
    omi = c(0, 0, 0, 0),
    cex.lab = 1)

## plot narwhal detections vs. ice cover
plot(narwhal ~ ice_cover_percent, data = mdl_data_pooled,
     pch = 16, cex = 1.3, col = clr,
     yaxt = "n", ylab = "Narwhal prob (%)", xlab = "Sea ice cover (%)")
axis(2, at = c(0,0.5,1), labels = c("0","50", "100"), las = 1)

# ## add 95% CI's
# lines(DOY, CI_upper, lwd = 1, col = "gray")
# lines(DOY, CI_lower, lwd = 1, col = "gray")
ice_cover_percent <- seq(0, 75)
polygon(c(ice_cover_percent, rev(ice_cover_percent)), c(CI_upper19, rev(CI_lower19)), border=NA,
        col = adjustcolor("orange",alpha.f=0.2))
polygon(c(ice_cover_percent, rev(ice_cover_percent)), c(CI_upper18, rev(CI_lower18)), border=NA,
        col = adjustcolor("gray",alpha.f=0.5))
## add model fit
lines(ice_cover_percent, p_hat18, lwd = 2, col="darkgray")
lines(ice_cover_percent, p_hat19, lwd = 2, col="orange")

## DOY ---------------------------------------
## plot narwhal prob vs DOY with confidence intervals
DOY <- seq(210,356)
## get fitted values
newdata18 <- data.frame(DOY = DOY,
                      ice_cover_percent = rep(mean(mdl_data_pooled$ice_cover_percent),length(DOY)),
                      year = as.factor(rep(2018,length(DOY))))
newdata19 <- data.frame(DOY = DOY,
                      ice_cover_percent = rep(mean(mdl_data_pooled$ice_cover_percent),length(DOY)),
                      year = as.factor(rep(2019,length(DOY))))

eta18 <- predict(mdl_best_open_water, newdata18)
eta19 <- predict(mdl_best_open_water, newdata19)
# transform out of logit space
p_hat18 <- 1 / (1 + exp(-eta18))
p_hat19 <- 1 / (1 + exp(-eta19))

## get the SE of the response
se18 <- predict(mdl_best_open_water, newdata18, type = "link", se.fit = TRUE)$se.fit
CI_upper18 <- 1 / (1 + exp(-(eta18 + 1.96*se18))) ## upper 95% CI
CI_lower18 <- 1 / (1 + exp(-(eta18 - 1.96*se18))) ## lower 95% CI

se19 <- predict(mdl_best_open_water, newdata19, type = "link", se.fit = TRUE)$se.fit
CI_upper19 <- 1 / (1 + exp(-(eta19 + 1.96*se19))) ## upper 95% CI
CI_lower19 <- 1 / (1 + exp(-(eta19 - 1.96*se19))) ## lower 95% CI

## 95% CI for beta_i based on profile likelihood
round(confint(mdl_best_open_water), 2)

## set plot area
par(mai = c(0.9, 0.9, 0.1, 0.1),
    omi = c(0, 0, 0, 0),
    cex.lab = 1)

## plot narwhal detections vs. ice cover
plot(narwhal ~ DOY, data = mdl_data_pooled,
     pch = 16, cex = 1.3, col = clr,
     yaxt = "n", ylab = "Narwhal prob (%)", xlab = "DOY")
axis(2, at = c(0,0.5,1), labels = c("0","50", "100"), las = 1)

# ## add 95% CI's
# lines(DOY, CI_upper, lwd = 1, col = "gray")
# lines(DOY, CI_lower, lwd = 1, col = "gray")
polygon(c(DOY, rev(DOY)), c(CI_upper19, rev(CI_lower19)), border=NA,
        col = adjustcolor("orange",alpha.f=0.2))
polygon(c(DOY, rev(DOY)), c(CI_upper18, rev(CI_lower18)), border=NA,
        col = adjustcolor("gray",alpha.f=0.5))
## add model fit
lines(DOY, p_hat18, lwd = 2, col="darkgray")
lines(DOY, p_hat19, lwd = 2, col="orange")

```


```{r global glm | whale occupation period}
## model using data for time period when whales were in the area
## dropped salt_deep and DOY since these were correlated with other vars

# make sure year is coded as a factor
mdl_data_pooled_whales <- mdl_data_pooled_whales %>% mutate_at(vars(year), list(factor))

mdl_global_whales <- glm(narwhal ~
                           temp_surf+I(temp_surf^2)+
                           temp_deep+I(temp_deep^2)+
                           salt_surf+I(salt_surf^2)+
                           ice_cover_percent+
                           runoff_racmo+
                           glacier_len_norm+
                           noise_4kHz+
                           year+
                           site,
                         data = mdl_data_pooled_whales,
                         family=binomial(link="logit"),
                         na.action="na.fail")

# print summary output
summary(mdl_global_whales)
faraway::sumary(mdl_global_whales)

dredge_mdl_whales <- dredge(mdl_global_whales)
dredge_mdl_whales
head(dredge_mdl_whales)

# export table to csv
write.csv(as.data.frame(dredge_mdl_whales), row.names = FALSE,
          here("analysis/csv_outputs/mdl_whale_occupancy_results.csv"))
```

```{r best model - whale occupancy}
mdl_best_whales <- glm(narwhal ~
                         ice_cover_percent+
                         # temp_deep+I(temp_deep^2)+ ## temp_deep was not in top model but in many of the runner up models
                         year,
                       data = mdl_data_pooled_whales,
                       family=binomial(link="logit"),
                       na.action="na.fail")
summary(mdl_best_whales)
faraway::sumary(mdl_best_whales)

# plot predictions
plot_model(mdl_best_whales)
plot_model(mdl_best_whales, type = 'pred',
           transform = "exp", show.data=TRUE)

```


Since we know the soundtrap detected more animals, I will run two separate models: one for all aural detections and one for soundtrap detections.

```{r run glm | separate aural and soundtrap models}
## model for only aural data (2018 and 2019)
## select only aural data
mdl_data_open_water_aural <- mdl_data_open_water %>% filter(device=='aural')

mdl_global_aural <- glm(narwhal ~
                    temp_surf+I(temp_surf^2)+
                    ice_cover_percent+
                    runoff_racmo+
                    velocity+
                    glacier_len_norm+
                    noise_4kHz+
                    year+
                    DOY+I(DOY^2)+
                    site,
                  data = mdl_data_open_water_aural,
                  family=binomial(link="logit"),
                  na.action="na.fail")

# print summary output
summary(mdl_global_aural)
faraway::sumary(mdl_global_aural)

dredge(mdl_global_aural)

# run best model for aural data
mdl_best_aural <- glm(narwhal ~
                    I(temp_surf^2)+
                    ice_cover_percent+
                    # salt_deep+I(salt_deep^2),
                    DOY+I(DOY^2),
                  data = mdl_data_open_water_aural,
                  family=binomial(link="logit"),
                  na.action="na.fail")
summary(mdl_best_aural)
plot_model(mdl_best_aural, type = 'pred',
           transform = "exp", show.data=TRUE)

## model for only soundtrap data (2019)
## select only aural data
mdl_data_open_water_soundtrap <- mdl_data_open_water %>% filter(device=='soundtrap')
mdl_data_whales_soundtrap <- mdl_data_whales %>% filter(device=='soundtrap')

mdl_global_soundtrap <- glm(narwhal ~
                    temp_surf+I(temp_surf^2)+
                    ice_cover_percent+
                    runoff_racmo+
                    glacier_length+
                    noise_4kHz+
                    # DOY+I(DOY^2)+
                    salt_deep+I(salt_deep^2)+
                    site,
                  data = mdl_data_open_water_soundtrap,
                  family=binomial(link="logit"),
                  na.action="na.fail")

# print summary output
summary(mdl_global_soundtrap)
faraway::sumary(mdl_global_soundtrap)

dredge(mdl_global_soundtrap)

# run best model for soundtrap data
mdl_best_soundtrap <- glm(narwhal ~
                    salt_deep+I(salt_deep^2),
                    # temp_deep+I(temp_deep^2),
                    # salt_surf+I(salt_surf^2), # does not show as good of a fit
                    # DOY+I(DOY^2),
                  data = mdl_data_open_water_soundtrap,
                  family=binomial(link="logit"),
                  na.action="na.fail")
summary(mdl_best_soundtrap)
faraway::sumary(mdl_best_soundtrap)
plot_model(mdl_best_soundtrap, type = 'pred',
           transform = "exp", show.data=TRUE)

```


Run separate glm for each glacier/site

```{r run glm | model for each site}
## Rink/Fisher ---------------------------------------------
mdl_data_rink <- mdl_data_pooled %>% filter(site=='rink')
mdl_rink <- glm(narwhal ~
                    temp_surf+I(temp_surf^2)+
                    ice_cover_percent+
                    runoff_racmo+
                    velocity+
                    glacier_len_norm+
                    noise_4kHz+
                    year+
                    DOY+I(DOY^2),
                  data = mdl_data_rink,
                  family=binomial(link="logit"),
                  na.action="na.fail")
summary(mdl_rink)
faraway::sumary(mdl_rink)
dredge(mdl_rink)

## Kong Oscar ----------------------------------------------
mdl_data_kong <- mdl_data_pooled %>% filter(site=='kong')
mdl_kong <- glm(narwhal ~
                    temp_surf+I(temp_surf^2)+
                    ice_cover_percent+
                    runoff_racmo+
                    velocity+
                    glacier_len_norm+
                    noise_4kHz+
                    year+
                    DOY+I(DOY^2),
                  data = mdl_data_kong,
                  family=binomial(link="logit"),
                  na.action="na.fail")
summary(mdl_kong)
dredge(mdl_kong)

## Sverdrup ------------------------------------------------
mdl_data_sver <- mdl_data_pooled %>% filter(site=='sver')
mdl_sver <- glm(narwhal ~
                    temp_surf+I(temp_surf^2)+
                    ice_cover_percent+
                    runoff_racmo+
                    velocity+
                    glacier_len_norm+
                    noise_4kHz+
                    DOY+I(DOY^2),
                  data = mdl_data_sver,
                  family=binomial(link="logit"),
                  na.action="na.fail")
summary(mdl_sver)
faraway::sumary(mdl_sver)
dredge(mdl_sver)

```


## Diagnostics

```{r diagnostics}
## plot deviance residuals------------------------------------------
## set plot area
par(mai = c(0.9, 0.9, 0.1, 0.1),
omi = c(0, 0, 0, 0),
cex.lab = 1)
## plot resids vs eta
binnedplot(fitted(mdl_global), 
           residuals(mdl_global), las = 1, pch = 16,
           ylab = "Residuals", xlab = "Fitted values",main = "")

binnedplot(fitted(mdl_global_whales), 
           residuals(mdl_global_whales), las = 1, pch = 16,
           ylab = "Residuals", xlab = "Fitted values",main = "")

## compute R^2
# deviances
# for open water period
nn <- length(mdl_data_open_water_pooled$time)
DM_mod <- mdl_global$deviance
D0_mod <- mdl_global$null.deviance
# R^2
(R2 <- (1 - exp((DM_mod - D0_mod) / nn)) / (1 - exp(-D0_mod / nn)))

# for period when whales were in the area
nn <- length(mdl_data_pooled_whales$time)
DM_mod <- mdl_global_whales$deviance
D0_mod <- mdl_global_whales$null.deviance
(R2 <- (1 - exp((DM_mod - D0_mod) / nn)) / (1 - exp(-D0_mod / nn)))

## leverages--------------------------------------------------------
## set plot area
par(mai = c(0.9, 1, 0.1, 0.1),
omi = c(0, 0, 0, 0),
cex.lab = 1)
levs <- hatvalues(mdl_global_whales)
## threshhold value
h_crit <- 2 * length(coef(mdl_global_whales)) / nn

## halfnormal plot
faraway::halfnorm(levs, las = 1, ylab = "")
text(0, 0.92*par("usr")[4], substitute(italic(h[crit]) == h_crit, 
                                       list(h_crit = h_crit)), pos = 4)
mtext(side = 2, text = "Sorted Data", line = 4)

## Cook's D---------------------------------------------------------
## set plot area
par(mai = c(0.9, 0.9, 0.1, 0.1),
omi = c(0, 0, 0, 0),
cex.lab = 1)
## halfnormal plot
CD <- cooks.distance(mdl_global_whales)
faraway::halfnorm(CD, las = 1, ylab = "")
mtext(side = 2, text = "Sorted Data", line = 3.5)

## H-L test with 8 groups
# A non-significant p value indicates that there is no evidence that the observed and expected frequencies differ (i.e., evidence of good fit).
library(generalhoslem)
generalhoslem::logitgof(obs = mdl_data_open_water_pooled$narwhal,
                        exp = fitted(mdl_global), g = 8)

generalhoslem::logitgof(obs = mdl_data_pooled_whales$narwhal,
                        exp = fitted(mdl_global_whales), g = 8)
```

```{r check autocorrelation, eval=FALSE}
acf(resid(mdl_global_open_water))
acf(resid(mdl_global_whales))
```

