---
title: "narwhal_model_data-cleaning"
author: "Marie Zahn"
date: '2023-01-26'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(tidyverse)
library(corrplot)
```


## Create master dataframe and check correlations

- Sea ice cover
- ice velocity
- temperature
- salinity
- runoff
- glacier length (terminus position: advance or retreat)
- ambient noise
- iceberg concentration

Additional glacier variables:
- glacier thickness
- grounding line
- glacier surface elevation
- glacier width
- runoff outlet depth
- mean runoff
- mean ice velocity
- mean solid ice discharge

```{r load data}
# use this line to list all filenames in specified directory
# files = list.files(data_dir='D:/Habitat_model_data/covariate_csv_files',pattern = '*.csv',recursive = TRUE)

# data stored on external harddrive -- update path here
data_dir <- 'H:/Habitat_model_data/covariate_csv_files/'

## load all data from csv files and add columns to indicate glacier/site and device (where applicable)
## then join all data into one dataframe per variable

# narwhal detections
narwhal_kong_AU18 <- read.csv(paste(data_dir,"narwhal_detections/Kong_Oscar_2018-2019_AuralM2_daily.csv",sep=''))%>% mutate(site="kong",device="aural") %>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_kong_AU19 <- read.csv(paste(data_dir,"narwhal_detections/Nuussuaq_KongOscar_AuralM2_daily.csv",sep=''))%>% mutate(site="kong",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_kong_ST19 <- read.csv(paste(data_dir,"narwhal_detections/Nuussuaq_KongOscar_ST_daily_wavTime.csv",sep=''))%>% mutate(site="kong",device="soundtrap")%>% dplyr::select(Time_UTC, narwhal,site,device) # wavTime file is actually UTC
narwhal_rink_AU18 <- read.csv(paste(data_dir,"narwhal_detections/Rink_Glacier_2018-19_AuralM2_daily.csv",sep=''))%>% mutate(site="rink",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_fish_AU19 <- read.csv(paste(data_dir,"narwhal_detections/Rink_Glacier_Fischer_AuralM2_daily.csv",sep=''))%>% mutate(site="rink",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_fish_ST19 <- read.csv(paste(data_dir,"narwhal_detections/Rink_Glacier_Fischer_ST_daily.csv",sep=''))%>% mutate(site="rink",device="soundtrap")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_sver_AU18 <- read.csv(paste(data_dir,"narwhal_detections/Sverdrup2018-19_AuralM2_daily.csv",sep=''))%>% mutate(site="sver",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)

narwhal_dets <- rbind(narwhal_kong_AU18, 
                      narwhal_kong_AU19, 
                      narwhal_kong_ST19, 
                      narwhal_rink_AU18, 
                      narwhal_fish_AU19, 
                      narwhal_fish_ST19, 
                      narwhal_sver_AU18) %>%
  mutate(time=as.character(strptime(as.Date(Time_UTC,format=c('%Y/%b/%d')),format='%Y-%m-%d'))) %>% dplyr::select(!Time_UTC)

# pool soundtrap and aural detections for 2019 data (i.e., take only one obs/day for each site)
narwhal_pooled <- narwhal_dets %>% group_by(site,time) %>% 
  filter(narwhal==max(narwhal),(row_number()==n()))

# ice velocity
kong_ice_vel <- read.csv(paste(data_dir,"glacier_ice_velocity/kongoscar_ice_velocity_1D.csv",sep=''))%>% mutate(site="kong")
rink_ice_vel <- read.csv(paste(data_dir,"glacier_ice_velocity/rink_ice_velocity_1D.csv",sep='')) %>% mutate(site="rink")
sver_ice_vel <- read.csv(paste(data_dir,"glacier_ice_velocity/sverdrup_ice_velocity_1D.csv",sep='')) %>% mutate(site="sver")
ice_vel<-rbind(kong_ice_vel,rink_ice_vel,sver_ice_vel)

# glacier length (= terminus position)
kong_glacier_length <- read.csv(paste(data_dir,"glacier_length/kongoscar_glacier_length_1D.csv",sep=''))%>% mutate(site="kong")
rink_glacier_length <- read.csv(paste(data_dir,"glacier_length/rink_glacier_length_1D.csv",sep=''))%>% mutate(site="rink")
sver_glacier_length <- read.csv(paste(data_dir,"glacier_length/sverdrup_glacier_length_1D.csv",sep=''))%>% mutate(site="sver")
glacier_len<-rbind(kong_glacier_length,rink_glacier_length,sver_glacier_length)

# runoff
# MAR
runoff_mar_kong <- read.csv(paste(data_dir,"runoff/kongoscar_runoff_mar_1D.csv",sep=''))%>% mutate(site="kong")
runoff_mar_rink <- read.csv(paste(data_dir,"runoff/rink_runoff_mar_1D.csv",sep=''))%>% mutate(site="rink")
runoff_mar_sver <- read.csv(paste(data_dir,"runoff/sverdrup_runoff_mar_1D.csv",sep=''))%>% mutate(site="sver")
runoff_mar <- rbind(runoff_mar_kong,runoff_mar_rink,runoff_mar_sver) %>% 
  rename(runoff_mar=runoff)

# RACMO
runoff_racmo_kong <- read.csv(paste(data_dir,"runoff/kongoscar_runoff_racmo_1D.csv",sep=''))%>% 
  mutate(site="kong")
runoff_racmo_rink <- read.csv(paste(data_dir,"runoff/rink_runoff_racmo_1D.csv",sep=''))%>% 
  mutate(site="rink")
runoff_racmo_sver <- read.csv(paste(data_dir,"runoff/sverdrup_runoff_racmo_1D.csv",sep=''))%>% 
  mutate(site="sver")
runoff_racmo <- rbind(runoff_racmo_kong,runoff_racmo_rink,runoff_racmo_sver) %>% 
  rename(runoff_racmo=runoff)

# sea ice
sea_ice_rink <- read.csv(paste(data_dir,"sea_ice/fisher_seaice_cover_1D.csv",sep=''))%>% mutate(site="rink")
sea_ice_kong <- read.csv(paste(data_dir,"sea_ice/kongoscar_seaice_cover_1D.csv",sep=''))%>% mutate(site="kong")
sea_ice_sver <- read.csv(paste(data_dir,"sea_ice/sverdrup_seaice_cover_1D.csv",sep=''))%>% mutate(site="sver")
sea_ice<-rbind(sea_ice_rink,sea_ice_kong,sea_ice_sver)

# salinity
salt_shallow_fish <- read.csv(paste(data_dir,"salinity/fisher_salt_120m_1D.csv",sep=''))%>% mutate(site="rink")
salt_deep_fish <- read.csv(paste(data_dir,"salinity/fisher_salt_220m_1D.csv",sep=''))%>% mutate(site="rink")
salt_shallow_kong <- read.csv(paste(data_dir,"salinity/kongoscar_salt_120m_1D.csv",sep=''))%>% mutate(site="kong") %>% mutate(time=as.character(as.Date(time,format='%m/%d/%Y')))
salt_deep_kong <- read.csv(paste(data_dir,"salinity/kongoscar_salt_220m_1D.csv",sep=''))%>% mutate(site="kong")
salt_shallow_rink <- read.csv(paste(data_dir,"salinity/rink_salt_120m_1D.csv",sep=''))%>% mutate(site="rink")
salt_deep_rink <- read.csv(paste(data_dir,"salinity/rink_salt_220m_1D.csv",sep=''))%>% mutate(site="rink")
salt_shallow_sver <- read.csv(paste(data_dir,"salinity/sverdrup_salt_120m_1D.csv",sep=''))%>% mutate(site="sver")
salt_deep_sver <- read.csv(paste(data_dir,"salinity/sverdrup_salt_220m_1D.csv",sep=''))%>% mutate(site="sver")
salt_shallow<-rbind(salt_shallow_fish,salt_shallow_kong,salt_shallow_rink,salt_shallow_sver) %>% rename(salt_shallow=salinity)
salt_deep<-rbind(salt_deep_fish,salt_deep_kong,salt_deep_rink,salt_deep_sver) %>% rename(salt_deep=salinity)

# temperature
temp_shallow_fish <- read.csv(paste(data_dir,"temperature/fisher_temp_120m_1D.csv",sep=''))%>% mutate(site="rink")
temp_deep_fish <- read.csv(paste(data_dir,"temperature/fisher_temp_220m_1D.csv",sep=''))%>% mutate(site="rink")
temp_shallow_kong <- read.csv(paste(data_dir,"temperature/kongoscar_temp_120m_1D.csv",sep=''))%>% mutate(site="kong")
temp_deep_kong <- read.csv(paste(data_dir,"temperature/kongoscar_temp_220m_1D.csv",sep=''))%>% mutate(site="kong")
temp_shallow_rink <- read.csv(paste(data_dir,"temperature/rink_temp_120m_1D.csv",sep=''))%>% mutate(site="rink")
temp_deep_rink <- read.csv(paste(data_dir,"temperature/rink_temp_220m_1D.csv",sep=''))%>% mutate(site="rink")
temp_shallow_sver <- read.csv(paste(data_dir,"temperature/sverdrup_temp_120m_1D.csv",sep=''))%>% mutate(site="sver")
temp_deep_sver <- read.csv(paste(data_dir,"temperature/sverdrup_temp_220m_1D.csv",sep=''))%>% mutate(site="sver")
temp_shallow <- rbind(temp_shallow_fish,temp_shallow_kong,temp_shallow_rink,temp_shallow_sver) %>% rename(temp_shallow=temperature)
temp_deep <- rbind(temp_deep_fish,temp_deep_kong,temp_deep_rink,temp_deep_sver) %>% rename(temp_deep=temperature)

# ## iceberg data
# ## calculate porportion of fjord ROI covered by icebergs (=percent area covered by icebergs)
# iceberg_rink_roi_area <- read.csv(paste(data_dir,"iceberg_concentration/rink_fjord_2017_2021.csv",sep='')) %>% select("Area_ROI_m2") %>% distinct()
# iceberg_kong_roi_area <- read.csv(paste(data_dir,"iceberg_concentration/kongoscar_fjord_2017_2021.csv",sep='')) %>% select("Area_ROI_m2") %>% distinct()
# iceberg_sver_roi_area <- read.csv(paste(data_dir,"iceberg_concentration/sverdrup_fjord_2017_2021.csv",sep='')) %>% select("Area_ROI_m2") %>% distinct()
# 
# iceberg_rink <- read.csv(paste(data_dir,"iceberg_concentration/rink_fjord_2017_2021.csv",sep='')) %>% select("AreaRF_m2","Date") %>% rename(time=Date) %>% group_by(time) %>% summarise(iceberg_m2 = sum(AreaRF_m2)) %>% mutate(iceberg_percent=iceberg_m2/iceberg_rink_roi_area$Area_ROI_m2*100) %>% mutate(site="rink")
# 
# iceberg_kong <- read.csv(paste(data_dir,"iceberg_concentration/kongoscar_fjord_2017_2021.csv",sep='')) %>% select("AreaRF_m2","Date") %>% rename(time=Date) %>% group_by(time) %>% summarise(iceberg_m2 = sum(AreaRF_m2)) %>% mutate(iceberg_percent=iceberg_m2/iceberg_kong_roi_area$Area_ROI_m2*100) %>% mutate(site="kong")
# 
# iceberg_sver <- read.csv(paste(data_dir,"iceberg_concentration/sverdrup_fjord_2017_2021.csv",sep='')) %>% select("AreaRF_m2","Date") %>% rename(time=Date) %>% group_by(time) %>% summarise(iceberg_m2 = sum(AreaRF_m2)) %>% mutate(iceberg_percent=iceberg_m2/iceberg_sver_roi_area$Area_ROI_m2*100) %>% mutate(site="sver")
# 
# iceberg<-rbind(iceberg_rink,iceberg_kong,iceberg_sver)
# 
# # export csv of iceberg dataset
# write.csv(iceberg, paste(data_dir,"iceberg_concentration/iceberg_conc.csv",sep=''))

# plot
# iceberg %>% ggplot(aes(x = time, y = iceberg_percent, group=site, color=site))+
  # geom_line()

## ambient noise
noise_kong_AU18 <- read.csv(paste(data_dir,"ambient_noise/Kong_Oscar_2018-2019_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="kong",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_kong_AU19 <- read.csv(paste(data_dir,"ambient_noise/Nuussuaq_KongOscar_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="kong",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_kong_ST19 <- read.csv(paste(data_dir,"ambient_noise/Nuussuaq_KongOscar_ST_UTC_daily_TOL.csv",sep=''))%>% mutate(site="kong",device='soundtrap') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_rink_AU18 <- read.csv(paste(data_dir,"ambient_noise/Rink_Glacier_2018-19_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="rink",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_rink_AU19 <- read.csv(paste(data_dir,"ambient_noise/Rink_Glacier_Fischer_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="rink",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_rink_ST19 <- read.csv(paste(data_dir,"ambient_noise/Rink_Glacier_Fischer_ST_UTC_daily_TOL.csv",sep=''))%>% mutate(site="rink",device='soundtrap') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_sver_AU18 <- read.csv(paste(data_dir,"ambient_noise/Sverdrup2018-19_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="sver",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)


noise <- rbind(noise_kong_AU18,
               noise_kong_AU19,
               noise_kong_ST19,
               noise_rink_AU18,
               noise_rink_AU19,
               noise_rink_ST19,
               noise_sver_AU18) %>%
  mutate(time=as.character(strptime(as.Date(Time,format=c('%d-%b-%Y')),format='%Y-%m-%d'))) %>% dplyr::select(!Time)

# table of additional glacier variables
glacier_vars_df <- read.csv(paste(data_dir,"narwhal_model_glacier_vars.csv",sep=''))

```

```{r tidy df}
## make data tidy and join all vars into one df

# ALL narwhal detections (aural + soundtrap)
model_vars_df <- runoff_racmo %>% 
  left_join(runoff_mar, by = c('time', 'site')) %>%
  left_join(salt_shallow, by = c('time', 'site')) %>% 
  left_join(salt_deep, by = c('time', 'site')) %>%
  left_join(temp_shallow, by = c('time', 'site')) %>%
  left_join(temp_deep, by = c('time', 'site')) %>%
  left_join(sea_ice, by = c('time', 'site')) %>% 
  left_join(ice_vel, by = c('time', 'site')) %>% 
  left_join(glacier_len, by = c('time', 'site')) %>%
  left_join(noise, by = c('time', 'site')) %>% 
  left_join(narwhal_dets, by = c('time', 'site','device')) %>% 
  mutate(DOY=as.numeric(strftime(as.Date(time),format='%j')), 
         year = strftime(as.Date(time),format='%Y')) 

# pooled narwhal detections (i.e., any positive detection from Aural or ST = 1)
narwhal_pooled_dets <- narwhal_pooled %>% dplyr::select(narwhal, site, device, time)
model_vars_pooled_df <- runoff_racmo %>% 
  left_join(runoff_mar, by = c('time', 'site')) %>%
  left_join(salt_shallow, by = c('time', 'site')) %>% 
  left_join(salt_deep, by = c('time', 'site')) %>%
  left_join(temp_shallow, by = c('time', 'site')) %>%
  left_join(temp_deep, by = c('time', 'site')) %>%
  left_join(sea_ice, by = c('time', 'site')) %>% 
  left_join(ice_vel, by = c('time', 'site')) %>% 
  left_join(glacier_len, by = c('time', 'site')) %>%
  left_join(noise, by = c('time', 'site')) %>% 
  left_join(narwhal_pooled_dets, by = c('time', 'site', 'device')) %>% 
  mutate(DOY=as.numeric(strftime(as.Date(time),format='%j')), 
         year = strftime(as.Date(time),format='%Y')) 

```

```{r pool data and get monthly sums}
mdl_data_pooled_group <- narwhal_pooled_dets %>% mutate(month=substr(time, start = 6, stop = 7)) %>% 
  mutate(year=substr(time, start = 1, stop = 4)) %>% 
  group_by(month, year) %>%
summarise(narwhal_sum = sum(narwhal))

mdl_data_pooled_group %>% filter(year==2018)
mdl_data_pooled_group %>% filter(year==2019)
```

```{r sanity check}
# check to make sure values are 
model_vars_df %>% filter(time=='2019-09-30')

# calculate how many days we have narwhal observations
days <- unique(model_vars_df %>% dplyr::select(time,narwhal) %>% drop_na() %>% dplyr::select(time))
length(days$time)
```

```{r select data for months Aug-Oct}
# only select period when whales were in the area
# filter all data before November 1st = DOY 305
mdl_data_2018 <- model_vars_df %>% filter(year==2018,DOY<305) %>% drop_na()
mdl_data_2019 <- model_vars_df %>% filter(year==2019,DOY<305 & DOY>100) %>% drop_na()

mdl_data <- rbind(mdl_data_2018,mdl_data_2019)

## select only days narwhals were present to find when the outlier point occurred in November 2019
model_vars_df %>% filter(narwhal==1,year==2019,site=="rink")

## we lose two detections because we didn't have temp/salt data for those days
model_vars_df %>% filter(narwhal==1,year==2018,site=="rink")
model_vars_df %>% filter(narwhal==1,year==2018,site=="kong")
model_vars_df %>% filter(narwhal==1,year==2018,site=="sver")

```

```{r export rdata file}
# export data as rdata file
# then I can open these files in separate script dedicated to model development
save(mdl_data, file="data-cleaning_R/narwhal_model_data_aug-oct.rdata")

```


## Summarize narwhal detections

```{r narwhal detection totals}
## pool soundtrap and aural detections for 2019 data (i.e., take only one obs/day for each site)
mdl_data_pooled_2019 <- model_vars_df %>% group_by(site,time) %>% 
  filter(year==2019, narwhal==max(narwhal),(row_number()==n()))

## identify how many detections from the same day occurred at more than one site
length(unique(model_vars_df_pooled$time)) # number of days (obs) for entire sampling period
obs_2018 <- model_vars_df_pooled %>% filter(year==2018)
obs_2018_jan <- model_vars_df_pooled %>% filter(year==2019,DOY<100)
obs_2019 <- model_vars_df_pooled %>% filter(year==2019,DOY>100)
obs_2020 <- model_vars_df_pooled %>% filter(year==2020)
length(unique(obs_2018$time))+length(unique(obs_2018_jan$time)) # number of days (obs) sampled in 2018
length(unique(obs_2019$time))+length(unique(obs_2020$time)) # number of days (obs) sampled in 2019
length(unique(mdl_data$time)) # number of days (obs) for open water period

sum(model_vars_df_pooled$narwhal) # total of 96 detections over two years at 3 sites
narwhal_dets_total <- model_vars_df_pooled %>% filter(narwhal==1) # number of days that narwhals were detected
length(unique(narwhal_dets_total$time)) # number of unique days when narwhals were identified (78 including outliers)

## determine how many days whales were seen at more than one glacier front
mult_det <- model_vars_df_pooled %>% dplyr::select(time, site, narwhal) %>% pivot_wider(names_from = site, values_from = narwhal) %>% mutate(rowsums = sum(kong,rink,sver,na.rm=TRUE)) %>% filter(rowsums>1)
length(mult_det$time)

## determine how many days whales were seen at more than one glacier front each year
doubles_2018 <- mult_det %>% filter(grepl("2018",time))
doubles_2019 <- mult_det %>% filter(grepl("2019",time))
length(doubles_2018$time)
length(doubles_2019$time)

## tally up total detections for each site for a given year
rink_2018 <- model_vars_df_pooled %>% filter(year==2018, site=='rink')
rink_2019 <- model_vars_df_pooled %>% filter(year==2019, site=='rink')
sum(rink_2018$narwhal)
sum(rink_2019$narwhal) # (add outlier that was removed)

kong_2018 <- model_vars_df_pooled %>% filter(year==2018, site=='kong')
kong_2019 <- model_vars_df_pooled %>% filter(year==2019, site=='kong')
sum(kong_2018$narwhal)
sum(kong_2019$narwhal)

sver_2018 <- model_vars_df_pooled %>% filter(year==2018, site=='sver')
sum(sver_2018$narwhal)

## extract last detection for each year
model_vars_df_pooled_2018 <- model_vars_df_pooled %>% filter(year==2018, narwhal==1)
model_vars_df_pooled_2019 <- model_vars_df_pooled %>% filter(year==2019, narwhal==1)
tail(model_vars_df_pooled_2018)
tail(model_vars_df_pooled_2019)
```

Out of the 78 days that narwhals were detected, 18 of those days were occasions when whales were detected at two locations on the same day. Whales were never detected at three locations on the same day in 2018.

Totals between each year/site

2018
* Rink: 14
* Kong Oscar: 9
* Sverdrup: 7

2019
* Rink: 32
* Kong Oscar: 34

## Compare detections between Aurals and SoundTraps in 2019

```{r aural vs soundtrap 2019}
## compare aural detections between years
aural_2018 <- model_vars_df %>% group_by(site) %>% filter(year==2018,device=='aural') %>% dplyr::select(time,narwhal) %>% rename(narwhal_aural=narwhal)
aural_2019 <- model_vars_df %>% group_by(site) %>% filter(year==2019,device=='aural') %>% dplyr::select(time,narwhal) %>% rename(narwhal_aural=narwhal)
sum(aural_2018$narwhal_aural)
sum(aural_2019$narwhal_aural)
## see whether there are any aural detections that were NOT detected by the soundtrap in 2019
sound_2019 <- model_vars_df %>% group_by(site) %>%  filter(year==2019,device=='soundtrap') %>% dplyr::select(time,narwhal)%>% rename(narwhal_sound=narwhal)

compare <- aural_2019 %>% group_by(site) %>%  left_join(sound_2019,by="time") %>% filter(site.x==site.y, narwhal_aural==1 & narwhal_sound==0)
compare # zero rows = no aural detections that were not picked up by soundtrap

## calculate the reverse - how many soundtrap detections were not matched in aurals
compare <- sound_2019 %>% group_by(site) %>%  left_join(aural_2019,by="time") %>% filter(site.x==site.y, narwhal_sound==1 & narwhal_aural==0)
compare

## check sums across years
sum(model_vars_df_pooled$narwhal)
model_vars_df %>% filter(year==2018) %>% dplyr::select(narwhal) %>% drop_na() %>% sum()
model_vars_df %>% filter(year==2019,device=='soundtrap') %>% dplyr::select(narwhal) %>% sum()

## double check comparison using different method
aural_2019_kong <- aural_2019 %>% filter(site=='kong') 
sound_2019_kong <- sound_2019 %>% filter(site=='kong') 
aural_2019_rink <- aural_2019 %>% filter(site=='rink') 
sound_2019_rink <- sound_2019 %>% filter(site=='rink') 

aural_2019_kong %>% left_join(sound_2019_kong,by="time") %>% filter(narwhal_aural==1&narwhal_sound==0)
aural_2019_rink %>% left_join(sound_2019_rink,by="time") %>% filter(narwhal_aural==1&narwhal_sound==0)

```

