---
title: "narwhal_model_data-cleaning"
author: "Marie Zahn"
date: '2023-01-26'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(tidyverse)
library(corrplot)
```


## Create master dataframe and check correlations

- Sea ice cover
- ice velocity
- temperature
- salinity
- runoff
- glacier length (terminus position: advance or retreat)
- ambient noise
- iceberg concentration

Additional glacier variables:
- glacier thickness
- grounding line
- glacier surface elevation
- glacier width
- runoff outlet depth
- mean runoff
- mean ice velocity
- mean solid ice discharge

```{r load data}
# use this line to list all filenames in specified directory
# files = list.files(data_dir='D:/Habitat_model_data/covariate_csv_files',pattern = '*.csv',recursive = TRUE)

# data stored on external harddrive -- update path here
data_dir <- 'H:/Habitat_model_data/covariate_csv_files/'

## load all data from csv files and add columns to indicate glacier/site and device (where applicable)
## then join all data into one dataframe per variable

# narwhal detections
narwhal_kong_AU18 <- read.csv(paste(data_dir,"narwhal_detections/Kong_Oscar_2018-2019_AuralM2_daily.csv",sep=''))%>% mutate(site="kong",device="aural") %>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_kong_AU19 <- read.csv(paste(data_dir,"narwhal_detections/Nuussuaq_KongOscar_AuralM2_daily.csv",sep=''))%>% mutate(site="kong",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_kong_ST19 <- read.csv(paste(data_dir,"narwhal_detections/Nuussuaq_KongOscar_ST_daily.csv",sep=''))%>% mutate(site="kong",device="soundtrap")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_rink_AU18 <- read.csv(paste(data_dir,"narwhal_detections/Rink_Glacier_2018-19_AuralM2_daily.csv",sep=''))%>% mutate(site="rink",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_fish_AU19 <- read.csv(paste(data_dir,"narwhal_detections/Rink_Glacier_Fischer_AuralM2_daily.csv",sep=''))%>% mutate(site="rink",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_fish_ST19 <- read.csv(paste(data_dir,"narwhal_detections/Rink_Glacier_Fischer_ST_daily.csv",sep=''))%>% mutate(site="rink",device="soundtrap")%>% dplyr::select(Time_UTC, narwhal,site,device)
narwhal_sver_AU18 <- read.csv(paste(data_dir,"narwhal_detections/Sverdrup2018-19_AuralM2_daily.csv",sep=''))%>% mutate(site="sver",device="aural")%>% dplyr::select(Time_UTC, narwhal,site,device)

narwhal_dets <- rbind(narwhal_kong_AU18, 
                      narwhal_kong_AU19, 
                      narwhal_kong_ST19, 
                      narwhal_rink_AU18, 
                      narwhal_fish_AU19, 
                      narwhal_fish_ST19, 
                      narwhal_sver_AU18) %>%
  mutate(time=as.character(strptime(as.Date(Time_UTC,format=c('%Y/%b/%d')),format='%Y-%m-%d'))) %>% dplyr::select(!Time_UTC)

# pool soundtrap and aural detections for 2019 data (i.e., take only one obs/day for each site)
narwhal_pooled <- narwhal_dets %>% group_by(site,time) %>% 
  filter(narwhal==max(narwhal),(row_number()==n()))

# ice velocity
kong_ice_vel <- read.csv(paste(data_dir,"glacier_ice_velocity/kongoscar_ice_velocity_1D.csv",sep=''))%>% mutate(site="kong")
rink_ice_vel <- read.csv(paste(data_dir,"glacier_ice_velocity/rink_ice_velocity_1D.csv",sep='')) %>% mutate(site="rink")
sver_ice_vel <- read.csv(paste(data_dir,"glacier_ice_velocity/sverdrup_ice_velocity_1D.csv",sep='')) %>% mutate(site="sver")
ice_vel<-rbind(kong_ice_vel,rink_ice_vel,sver_ice_vel)

# glacier length (= terminus position)
kong_glacier_length <- read.csv(paste(data_dir,"glacier_length/kongoscar_glacier_length_1D.csv",sep=''))%>% mutate(site="kong")
rink_glacier_length <- read.csv(paste(data_dir,"glacier_length/rink_glacier_length_1D.csv",sep=''))%>% mutate(site="rink")
sver_glacier_length <- read.csv(paste(data_dir,"glacier_length/sverdrup_glacier_length_1D.csv",sep=''))%>% mutate(site="sver")
glacier_len<-rbind(kong_glacier_length,rink_glacier_length,sver_glacier_length)

# runoff
# MAR
runoff_mar_kong <- read.csv(paste(data_dir,"runoff/kongoscar_runoff_mar_1D.csv",sep=''))%>% mutate(site="kong")
runoff_mar_rink <- read.csv(paste(data_dir,"runoff/rink_runoff_mar_1D.csv",sep=''))%>% mutate(site="rink")
runoff_mar_sver <- read.csv(paste(data_dir,"runoff/sverdrup_runoff_mar_1D.csv",sep=''))%>% mutate(site="sver")
runoff_mar <- rbind(runoff_mar_kong,runoff_mar_rink,runoff_mar_sver) %>% 
  rename(runoff_mar=runoff)

# RACMO
runoff_racmo_kong <- read.csv(paste(data_dir,"runoff/kongoscar_runoff_racmo_1D.csv",sep=''))%>% 
  mutate(site="kong")
runoff_racmo_rink <- read.csv(paste(data_dir,"runoff/rink_runoff_racmo_1D.csv",sep=''))%>% 
  mutate(site="rink")
runoff_racmo_sver <- read.csv(paste(data_dir,"runoff/sverdrup_runoff_racmo_1D.csv",sep=''))%>% 
  mutate(site="sver")
runoff_racmo <- rbind(runoff_racmo_kong,runoff_racmo_rink,runoff_racmo_sver) %>% 
  rename(runoff_racmo=runoff)

# sea ice
sea_ice_rink <- read.csv(paste(data_dir,"sea_ice/fisher_seaice_cover_1D.csv",sep=''))%>% mutate(site="rink")
sea_ice_kong <- read.csv(paste(data_dir,"sea_ice/kongoscar_seaice_cover_1D.csv",sep=''))%>% mutate(site="kong")
sea_ice_sver <- read.csv(paste(data_dir,"sea_ice/sverdrup_seaice_cover_1D.csv",sep=''))%>% mutate(site="sver")
sea_ice<-rbind(sea_ice_rink,sea_ice_kong,sea_ice_sver)

# salinity
salt_shallow_fish <- read.csv(paste(data_dir,"salinity/fisher_salt_120m_1D.csv",sep=''))%>% mutate(site="rink")
salt_deep_fish <- read.csv(paste(data_dir,"salinity/fisher_salt_220m_1D.csv",sep=''))%>% mutate(site="rink")
salt_shallow_kong <- read.csv(paste(data_dir,"salinity/kongoscar_salt_120m_1D.csv",sep=''))%>% mutate(site="kong") %>% mutate(time=as.character(as.Date(time,format='%m/%d/%Y')))
salt_deep_kong <- read.csv(paste(data_dir,"salinity/kongoscar_salt_220m_1D.csv",sep=''))%>% mutate(site="kong")
salt_shallow_rink <- read.csv(paste(data_dir,"salinity/rink_salt_120m_1D.csv",sep=''))%>% mutate(site="rink")
salt_deep_rink <- read.csv(paste(data_dir,"salinity/rink_salt_220m_1D.csv",sep=''))%>% mutate(site="rink")
salt_shallow_sver <- read.csv(paste(data_dir,"salinity/sverdrup_salt_120m_1D.csv",sep=''))%>% mutate(site="sver")
salt_deep_sver <- read.csv(paste(data_dir,"salinity/sverdrup_salt_220m_1D.csv",sep=''))%>% mutate(site="sver")
salt_shallow<-rbind(salt_shallow_fish,salt_shallow_kong,salt_shallow_rink,salt_shallow_sver) %>% rename(salt_shallow=salinity)
salt_deep<-rbind(salt_deep_fish,salt_deep_kong,salt_deep_rink,salt_deep_sver) %>% rename(salt_deep=salinity)

# temperature
temp_shallow_fish <- read.csv(paste(data_dir,"temperature/fisher_temp_120m_1D.csv",sep=''))%>% mutate(site="rink")
temp_deep_fish <- read.csv(paste(data_dir,"temperature/fisher_temp_220m_1D.csv",sep=''))%>% mutate(site="rink")
temp_shallow_kong <- read.csv(paste(data_dir,"temperature/kongoscar_temp_120m_1D.csv",sep=''))%>% mutate(site="kong")
temp_deep_kong <- read.csv(paste(data_dir,"temperature/kongoscar_temp_220m_1D.csv",sep=''))%>% mutate(site="kong")
temp_shallow_rink <- read.csv(paste(data_dir,"temperature/rink_temp_120m_1D.csv",sep=''))%>% mutate(site="rink")
temp_deep_rink <- read.csv(paste(data_dir,"temperature/rink_temp_220m_1D.csv",sep=''))%>% mutate(site="rink")
temp_shallow_sver <- read.csv(paste(data_dir,"temperature/sverdrup_temp_120m_1D.csv",sep=''))%>% mutate(site="sver")
temp_deep_sver <- read.csv(paste(data_dir,"temperature/sverdrup_temp_220m_1D.csv",sep=''))%>% mutate(site="sver")
temp_shallow <- rbind(temp_shallow_fish,temp_shallow_kong,temp_shallow_rink,temp_shallow_sver) %>% rename(temp_shallow=temperature)
temp_deep <- rbind(temp_deep_fish,temp_deep_kong,temp_deep_rink,temp_deep_sver) %>% rename(temp_deep=temperature)

# ## iceberg data
# ## calculate porportion of fjord ROI covered by icebergs (=percent area covered by icebergs)
# iceberg_rink_roi_area <- read.csv(paste(data_dir,"iceberg_concentration/rink_fjord_2017_2021.csv",sep='')) %>% select("Area_ROI_m2") %>% distinct()
# iceberg_kong_roi_area <- read.csv(paste(data_dir,"iceberg_concentration/kongoscar_fjord_2017_2021.csv",sep='')) %>% select("Area_ROI_m2") %>% distinct()
# iceberg_sver_roi_area <- read.csv(paste(data_dir,"iceberg_concentration/sverdrup_fjord_2017_2021.csv",sep='')) %>% select("Area_ROI_m2") %>% distinct()
# 
# iceberg_rink <- read.csv(paste(data_dir,"iceberg_concentration/rink_fjord_2017_2021.csv",sep='')) %>% select("AreaRF_m2","Date") %>% rename(time=Date) %>% group_by(time) %>% summarise(iceberg_m2 = sum(AreaRF_m2)) %>% mutate(iceberg_percent=iceberg_m2/iceberg_rink_roi_area$Area_ROI_m2*100) %>% mutate(site="rink")
# 
# iceberg_kong <- read.csv(paste(data_dir,"iceberg_concentration/kongoscar_fjord_2017_2021.csv",sep='')) %>% select("AreaRF_m2","Date") %>% rename(time=Date) %>% group_by(time) %>% summarise(iceberg_m2 = sum(AreaRF_m2)) %>% mutate(iceberg_percent=iceberg_m2/iceberg_kong_roi_area$Area_ROI_m2*100) %>% mutate(site="kong")
# 
# iceberg_sver <- read.csv(paste(data_dir,"iceberg_concentration/sverdrup_fjord_2017_2021.csv",sep='')) %>% select("AreaRF_m2","Date") %>% rename(time=Date) %>% group_by(time) %>% summarise(iceberg_m2 = sum(AreaRF_m2)) %>% mutate(iceberg_percent=iceberg_m2/iceberg_sver_roi_area$Area_ROI_m2*100) %>% mutate(site="sver")
# 
# iceberg<-rbind(iceberg_rink,iceberg_kong,iceberg_sver)
# 
# # export csv of iceberg dataset
# write.csv(iceberg, paste(data_dir,"iceberg_concentration/iceberg_conc.csv",sep=''))

# plot
# iceberg %>% ggplot(aes(x = time, y = iceberg_percent, group=site, color=site))+
  # geom_line()

## ambient noise
noise_kong_AU18 <- read.csv(paste(data_dir,"ambient_noise/Kong_Oscar_2018-2019_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="kong",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_kong_AU19 <- read.csv(paste(data_dir,"ambient_noise/Nuussuaq_KongOscar_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="kong",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_kong_ST19 <- read.csv(paste(data_dir,"ambient_noise/Nuussuaq_KongOscar_ST_UTC_daily_TOL.csv",sep=''))%>% mutate(site="kong",device='soundtrap') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_rink_AU18 <- read.csv(paste(data_dir,"ambient_noise/Rink_Glacier_2018-19_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="rink",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_rink_AU19 <- read.csv(paste(data_dir,"ambient_noise/Rink_Glacier_Fischer_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="rink",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_rink_ST19 <- read.csv(paste(data_dir,"ambient_noise/Rink_Glacier_Fischer_ST_UTC_daily_TOL.csv",sep=''))%>% mutate(site="rink",device='soundtrap') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)
noise_sver_AU18 <- read.csv(paste(data_dir,"ambient_noise/Sverdrup2018-19_AuralM2_daily_TOL.csv",sep=''))%>% mutate(site="sver",device='aural') %>% 
  dplyr::select(Time,X100Hz,X2000Hz,X4000Hz,site,device) %>% 
  rename(noise_100Hz=X100Hz, noise_2kHz=X2000Hz, noise_4kHz=X4000Hz)


noise <- rbind(noise_kong_AU18,
               noise_kong_AU19,
               noise_kong_ST19,
               noise_rink_AU18,
               noise_rink_AU19,
               noise_rink_ST19,
               noise_sver_AU18) %>%
  mutate(time=as.character(strptime(as.Date(Time,format=c('%d-%b-%Y')),format='%Y-%m-%d'))) %>% dplyr::select(!Time)

# table of additional glacier variables
glacier_vars_df <- read.csv(paste(data_dir,"narwhal_model_glacier_vars.csv",sep=''))

```

```{r tidy df}
## make data tidy and join all vars into one df

# ALL narwhal detections (aural + soundtrap)
model_vars_df <- runoff_racmo %>% 
  left_join(runoff_mar, by = c('time', 'site')) %>%
  left_join(salt_shallow, by = c('time', 'site')) %>% 
  left_join(salt_deep, by = c('time', 'site')) %>%
  left_join(temp_shallow, by = c('time', 'site')) %>%
  left_join(temp_deep, by = c('time', 'site')) %>%
  left_join(sea_ice, by = c('time', 'site')) %>% 
  left_join(ice_vel, by = c('time', 'site')) %>% 
  left_join(glacier_len, by = c('time', 'site')) %>%
  left_join(noise, by = c('time', 'site')) %>% 
  left_join(narwhal_dets, by = c('time', 'site','device')) %>% 
  mutate(DOY=as.numeric(strftime(as.Date(time),format='%j')), 
         year = strftime(as.Date(time),format='%Y')) 

# pooled narwhal detections (i.e., any positive detection from Aural or ST = 1)
narwhal_pooled_dets <- narwhal_pooled %>% dplyr::select(narwhal, site, device, time)
model_vars_pooled_df <- runoff_racmo %>% 
  left_join(runoff_mar, by = c('time', 'site')) %>%
  left_join(salt_shallow, by = c('time', 'site')) %>% 
  left_join(salt_deep, by = c('time', 'site')) %>%
  left_join(temp_shallow, by = c('time', 'site')) %>%
  left_join(temp_deep, by = c('time', 'site')) %>%
  left_join(sea_ice, by = c('time', 'site')) %>% 
  left_join(ice_vel, by = c('time', 'site')) %>% 
  left_join(glacier_len, by = c('time', 'site')) %>%
  left_join(noise, by = c('time', 'site')) %>% 
  left_join(narwhal_pooled_dets, by = c('time', 'site', 'device')) %>% 
  mutate(DOY=as.numeric(strftime(as.Date(time),format='%j')), 
         year = strftime(as.Date(time),format='%Y')) 

```

```{r sanity check}
# check to make sure values are 
model_vars_df %>% filter(time=='2019-09-30')

# calculate how many days we have narwhal observations
days <- unique(model_vars_df %>% dplyr::select(time,narwhal) %>% drop_na() %>% dplyr::select(time))
length(days$time)
```

```{r select data for months Aug-Oct}
# only select period when whales were in the area
# filter all data before November 1st = DOY 305
mdl_data_2018 <- model_vars_df %>% filter(year==2018,DOY<305) %>% drop_na()
mdl_data_2019 <- model_vars_df %>% filter(year==2019,DOY<305 & DOY>100) %>% drop_na()

mdl_data <- rbind(mdl_data_2018,mdl_data_2019)

## select only days narwhals were present to find when the outlier point occurred in November 2019
model_vars_df %>% filter(narwhal==1,year==2019,site=="rink")

```


```{r export rdata file}
# export data as rdata file
# then I can open these files in separate script dedicated to model development
save(mdl_data, file="data-cleaning_R/narwhal_model_data_aug-oct.rdata")

```

